---
title: "R Notebook"
output: html_notebook
---



```{r setup, include=FALSE}
library(tidyverse)
library(psych)
# require(devtools)
# install_version("mnormt", version = "1.5-7", repos = "http://cran.us.r-project.org")
```

# Read processed data
** These columns are removed because of 0.70 correlation with other columns**
- Gain index
- urban_percent
** These columns are removed because they are missing more than 30% data **
- Percent population internally displaced due  to conflict and violence
- Percent population internally displaced by a disaster


```{r}
(mod_comb_2017 <- read_csv("../Processed_data/combined_final_data_frame.csv") %>%
   dplyr::select(-gain_index_value,-percent_dis_conflict_population,-percent_dis_disaster_population,-urban_percent))
```


** Higher MCV1_percent_coverage is better **
** Higher HDI index is better **
** Higher climate risk index is better **
** Higher overall_vaccine_confidence_percent is better **
** Lower peace index value is better **
** Lower percent_refugee_origin_population is better **
** Lower measles_incidence_rate_per_100_000 is better **

```{r}
 describe(mod_comb_2017) %>%
  as_tibble(rownames = "id") %>%
  filter(!is.nan(mean)) %>%
  select(id,min,max)
```

# Final vaccine risk index ranking with climate change
```{r}
(
  mod_comb_2017 %>%
    dplyr::select(-Year) %>%
    mutate(
      measles_incidence_rate_per_100_000 = scales::rescale(measles_incidence_rate_per_100_000, to =
                                                             c(0, 1))
    ) %>%
    mutate(climate_risk_index = scales::rescale(climate_risk_index, to = c(0, 1))) %>%
    
    mutate(peace_index_score = scales::rescale(peace_index_score, to =
                                                 c(0, 1))) %>%
    
    mutate_at(
      vars(
        MCV1_percent_coverage,
        percent_refugee_origin_population
      ),
      funs(. / 100)
    ) %>%
    
    # Convert negative factors into numbers where higher values indicates positive change
    # Since numbers are scaled from 0-1, we can subtract the values from 1 to get the values from 0 to 1  so that implies lower values into better ranking
    
    mutate_at (
      vars(
        MCV1_percent_coverage,
        climate_risk_index,
        HDI_index_value,
        overall_vaccine_confidence_percent
      ),
      funs(1 - .)
    )   %>%
    
    mutate (
      na_count = dplyr::select(
        .,
        MCV1_percent_coverage,
        overall_vaccine_confidence_percent,
        climate_risk_index,
        HDI_index_value,
        percent_refugee_origin_population,
        measles_incidence_rate_per_100_000,
        peace_index_score
      ) %>% pmap_dbl(~ sum(is.na(c(
        ...
      ))))
    ) %>%
    
    filter(na_count < 2) %>% #Remove countries missing more than 1 column of data
    
    dplyr::select(
      Continent,
      ISO_code,
      Cname,
      MCV1_percent_coverage,
      overall_vaccine_confidence_percent,
      climate_risk_index,
      HDI_index_value,
      percent_refugee_origin_population,
      measles_incidence_rate_per_100_000,
      peace_index_score
    )  %>%
    write_csv(
      .,
      "../Processed_data/tentative_vaccine_risk_data.csv"
    )
)
```